### 2017-01-23 chapters/ch2-bottom-up.tex
Is the traditional Byzantine Agreement too strict?

The Swirlds~\cite{baird2016swirlds} algorithm is a departure from the classical
BFT algorithms. Nodes use ``gossip about gossip'' to determine whether a
majority of the population has heard about a particular transaction. So
agreement is achieved not after one execution of the algorithm but after the
nodes have eventually exchanged enough messages---we call this ``eventual
consensus''. The downside of Swirlds is that they assume static membership, thus
it is not permissionless. Swirlds may be applicable to our design where we can
use our group formation technique to select the validators and run Swirlds on
top (assuming Swirlds scales well). This eleminates the need for the recursive
(bottom-up) phase.

% TODO Swirlds need good scalability and convergence properties to be useful for
% us. With our bootstrapping, perhaps we can infer the member count in Swirlds
% from its history!

@@

>>>>

@@

\section{Is the traditional Byzantine Agreement too strict?}
If nodes are in agreement, they share the same state machine. But is this too
strict for our use case?

Suppose there is set of transactions $A \rightarrow B \rightarrow C \rightarrow
D$ that depend on each other. Then half of the nodes agree on a subset of those
transactions, i.e. $\{A, B, C\}$. Another half agree on all four transactions.
This is not Byzantine agreement in the traditional sense, because there does not
exist a majority that share the same state machine. But indirectly they all
agree on the set intersection. Can we build a model that support this type of
consensus?
